Topological Equivariant Artist Model

### 2021-03-24 slides/main.tex
transition into code

\begin{frame}{Model data as a fiber bundle \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
    A fiber bundle is a tuple $(\dtotal,\,\dbase,\,\pi ,\,\dfiber)$ defined by the map $\pi$
    \begin{equation}
        \label{eq:fiber_bundle}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[r, "\pi"] \& \dbase
        \end{tikzcd}
    \end{equation}
    \begin{description}
        \item[base space \dbase] coordinate system agnostic \textit{keys}
        \item[fiber space \dfiber] contains \textit{values}    
    \end{description}
    \begin{block}{Structural \textit{keys} with associated \textit{values} \cite{munznerVisualizationAnalysisDesign2014}}
        \begin{figure}
            \includegraphics[width=1\textwidth]{figures/intro/munzner_datatypes.png}
            \caption{Figure 2.8 in Munzner's Visualization Analysis and Design\cite{munznerVisualizationAnalysisDesign2014}}
        \end{figure}
    \end{block}

@@

\begin{frame}{}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/base.png}
        \label{fig:base_example}
    \end{figure}
    \begin{block}{\dbase\ is an indexing space regardless of components}
        \begin{equation}
            \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
        \end{equation}
    \end{block}
\end{frame}

\begin{frame}{Data are sections\dsection\ on the \dtotal}
    For any fiber bundle, there exists a map
    \begin{equation}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[d, "\pi"'] \\
                              \& \dbase \arrow[u, "\dsection"', bend right]
        \end{tikzcd}
    \end{equation}
     s.t. $\pi(\dsection(\dbasepoint)) = \dbasepoint$. $\Gamma(\dtotal)$ is the set of all global sections.
     \begin{figure}[H]
        \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
        \label{fig:data_sections}
    \end{figure}
\end{frame}

@@

\begin{frame}{Equivalance class of artists}
    An approximation of the equivalence class of artists $\vartist^{\prime}$
    \begin{equation}
    \vartist \in \vartist^{\prime}: \vartist_{1} \equiv \vartist_{2}
    \end{equation}
    roughly treats two artists as equivalent if they 
    \begin{itemize}
        \item act on the same visual bundle \vtotal
        \item have the same assembly function \vmark 
        \item have the same continuity map \vindex 
    \end{itemize}
\end{frame}

\section{Prototype}
\subsection{Artist}
\begin{frame}[fragile]{Artist}
    \begin{minted}{python}
        class ArtistClass(matplotlib.artist.Artist):
            def __init__(self, data, transforms, *args, **kwargs):
                # properties that are specific to the graphic
                self.data = data 
                self.transforms = transforms
                super().__init__(*args, **kwargs)

@@

\begin{frame}[fragile]{Artists: Scatter \& Line}
\begin{minted}{python}
class Point(mcollections.Collection):
    def assemble(self, x, y, s, facecolors='C0' ):
        # construct geometries of the circle glyphs in visual coordinates
        self._paths = [mpath.Path.circle(center=(xi,yi), radius=si) 
                    for (xi, yi, si) in zip(x, y, s)] 
        # set attributes of glyphs, these are vectorized 
        # circles and facecolors are lists of the same size
        self.set_facecolors(facecolors)
\end{minted} 
    \pause
\begin{minted}{python}
class Line(mcollections.LineCollection):
    def assemble(self, x, y, color='C0'):
        #assemble line marks as set of segments 
        segments = [np.vstack((vx, vy)).T for vx, vy in zip(x, y)]
        self.set_segments(segments)
        self.set_color(color)
\end{minted}
\end{frame}

\begin{frame}[fragile]{Visual Transformations}
\begin{minted}{python}
cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
transforms = {'x': {'name': 'v4', 'encoder': lambda x: x},
                'y': {'name': 'v2', 'encoder': lambda x: x},
                'facecolors': {'name':'v3', 'encoder': cmap}, 
                's':{'name': None , 
                    'encoder': lambda _: itertools.repeat(.02)}}
\end{minted}
    \begin{itemize}
        \item \mintinline{python}{lambda x: x} is identity \vchannel\
        \item \mintinline{python}|{'name':None}| map into \vfiber\ without corresponding \dsection\
        \item \mintinline{python}{color.Categorical} is custom \vchannel
    \end{itemize}
\end{frame}

\begin{frame}[fragile]{Custom Complex \vchannel}
\begin{minted}{python}
class Categorical:
    def __init__(self, mapping):
        # check that the conversion is to valid colors
        assert(mcolors.is_color_like(color) for color in mapping.values())
        self._mapping = mapping

    def __call__(self, value):
        # convert value to a color
        return [mcolors.to_rgba(self._mapping[v]) for v in values]
\end{minted}
That we can test for action equivariance
\begin{minted}{python}
def test_nominal(values, encoder):
    m1 = list(zip(values, encoder(values)))
    random.shuffle(values)
    m2 = list(zip(values, encoder(values)))
    assert sorted(m1) == sorted(m2)
\end{minted}
\end{frame}
\subsection{Data}
\begin{frame}[fragile]{Fiber Bundle}
\begin{minted}{python}
@dataclass
class FiberBundle:
"""
Attributes
----------
K: {'tables': []}
F: {variable name: type}
"""
    K: dict 
    F: dict
\end{minted}
\end{frame}

\begin{frame}[fragile]{Discrete Connectivity}
\begin{minted}{python}
class VertexSimplex: #maybe change name to something else
    """Fiberbundle is consistent across all sections
    """
    FB = FiberBundle({'tables': ['vertex']},  
            {'v1': float, 'v2': str, 'v3': float})

    def __init__(self, sid = 45, size=1000, max_key=10**10):
        # create random list of keys
    def tau(self, k):
        # e1 is sampled from F1, e2 from F2, etc...
        return (k, (e1, e2, e3, e4))
\end{minted}
\end{frame}

\begin{frame}[fragile]{1D Continuous Connectivity}
\begin{minted}{python}
class EdgeSimplex: 
    FB = FiberBundle({'tables': ['vertex','edge']},
                         {'x' : float, 'y':  float, 
                         'color':mtypes.Color()}})
    def __init__(self, num_edges=4, num_samples=1000): 
        self.keys = range(num_edge) #edge id
        self.distances = np.linspace(0,1, num_samples)
        # half generlized representation of arcs on a circle
        self.angle_samples = np.linspace(0, 2*np.pi, len(self.keys)+1)
    @staticmethod
    def _color(edge):
        return ['red','orange', 'green','blue'][edge%4]
    @staticmethod
    def _xy(edge, distances, start=0, end=2*np.pi):
        # start and end are parameterizations b/c really there is 
        angles = (distances *(end-start)) + start
        return np.cos(angles), np.sin(angles)
    def tau(self, k): #will fix location on page on revision
        x, y = self._xy(k, self.distances, 
                        self.angle_samples[k], self.angle_samples[k+1]) 
        color = self._color(k) 
        return (k, (x, y, color))
\end{minted}
\end{frame}

@@

\section{Conclusion}
\begin{frame}{Summary}
    \begin{itemize}
        \item structure preserving maps from data to visual representation:
            \begin{itemize}
                \item data and graphics have equivalent continuity
                \item properties are equivariant under monoid actions
            \end{itemize}  
        \item fiber bundles with a schema are structure rich abstractions of 
            \begin{itemize}
                \item topologically complex heterogenous data 
                \item target display spaces
            \end{itemize}
        \item model can be iteratively integrated into existing Matplotlib architecture
    \end{itemize}
\end{frame}

@@

\begin{frame}{Glyph}
    The glyph is the graphic generated by $\vmark(\gbase_{\dbasepathpoint})$ where the path connected components $\dbasepath \subset \dbase$ are defined 
    \begin{equation}
    \dbasepath = \{\dbasepathpoint \in \dbase \text{ s. t. } \exists \gamma \text{ s.t. } \gamma(0)=\dbasepoint \text{ and }\gamma(1)=\dbasepathpoint\}
    \end{equation}
    such that the  path $\gamma$ from \dbasepoint\ to \dbasepathpoint\ is a continuous function from the interval [0,1] and $\gbase_{\dbasepathpoint}$ is the region
    \begin{equation}
        \begin{tikzcd}[ampersand replacement=\&]
            \gtotal \arrow[r, shift left] \& \gbase_\dbasepathpoint \arrow[rr, "\vindex(\gbasepoint)", shift left] \arrow[l, "\gsection(\gbase_\dbasepathpoint)"] \& \& \dbasepath_{\dbasepoint} \arrow[ll, "\vindex^{-1}(\dbasepath)"]
            \end{tikzcd}
        \label{eq:mark}
    \end{equation}
    such that the glyph is differentiable, in keeping with Ziemkiewicz and Kosara's description of a glyph\cite{ziemkiewiczEmbeddingInformationVisualization2009}.
\end{frame}

>>>>

\begin{frame}{Model data as a fiber bundle \cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}}
    A fiber bundle is a tuple $(\dtotal,\,\dbase,\,\pi ,\,\dfiber)$ defined by the map $\pi$
    \begin{equation}
        \label{eq:fiber_bundle}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[r, "\pi"] \& \dbase
        \end{tikzcd}
    \end{equation}
    \begin{description}
        \item[base space \dbase] continuity
        \item[fiber space \dfiber] variable characteristics
        \item[total space \dtotal] dataset topology
    \end{description}
\end{frame}

@@

\begin{frame}{Structural \textit{keys} with associated \textit{values} \cite{munznerVisualizationAnalysisDesign2014}}
    \begin{figure}[H]
        \includegraphics[width=1\textwidth]{figures/math/base.png}
        \label{fig:base_example}
    \end{figure}

    \begin{block}{\dbase\ is an indexing space independent of component semantics}
        \begin{equation}
            \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
        \end{equation}
    \end{block}
\end{frame}

\begin{frame}{Data are sections \dsection\ on \dtotal}
    For any fiber bundle, there exists a map
    \begin{equation}
        \begin{tikzcd}[ampersand replacement=\&]
            \dfiber \arrow[r, hook] \& \dtotal \arrow[d, "\pi"'] \\
                              \& \dbase \arrow[u, "\dsection"', bend right]
        \end{tikzcd}
    \end{equation}
     s.t. $\pi(\dsection(\dbasepoint)) = \dbasepoint$. $\Gamma(\dtotal)$ is the set of all global sections.
     \begin{figure}[H]
        \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
        \label{fig:data_sections}
    \end{figure}
\end{frame}

@@

\begin{frame}{Equivalance class of artists}
    An approximation of the equivalence class of artists $\vartist^{\prime}$
    \begin{equation}
    \vartist \in \vartist^{\prime}: \vartist_{1} \equiv \vartist_{2}
    \end{equation}
    roughly treats two artists as equivalent if they 
    \begin{itemize}
        \item act on the same visual bundle \vtotal
        \item have the same assembly function \vmark 
        \item have the same continuity map \vindex 
    \end{itemize}
\end{frame}
\section{Prototype}

\begin{frame}{TEAM driven rearchitecture of Matplotlib}
    \begin{itemize}
        \item fiber bundle with schema is a structure rich abstraction
        \begin{itemize}
            \item topologically complex heterogenous data 
            \item target display spaces
        \end{itemize}
        \item structure preserving maps from data to visual:
            \begin{itemize}
                \item data and graphics have equivalent continuity
                \item properties are equivariant under monoid actions
            \end{itemize}  
        \item composable into complex visualizations
    \end{itemize}
\end{frame}

\subsection{Visual Transforms}
\begin{frame}[fragile]{Visual Transformations}
    \begin{minted}{python}
    cmap =  color.Categorical({'true':'deeppink', 'false':'deepskyblue'})
    transforms = {'x': {'name': 'v4', 'encoder': lambda x: x},
                    'y': {'name': 'v2', 'encoder': lambda x: x},
                    'facecolors': {'name':'v3', 'encoder': cmap}, 
                    's':{'name': None , 
                        'encoder': lambda _: itertools.repeat(.02)}}
    \end{minted}
        \begin{itemize}
            \item \mintinline{python}{lambda x: x} is identity \vchannel\
            \item \mintinline{python}|{'name':None}| map into \vfiber\ without corresponding \dsection\
            \item \mintinline{python}{color.Categorical} is custom \vchannel
        \end{itemize}
    \end{frame}

    \begin{frame}[fragile]{Complex \vchannel}
    \begin{minted}{python}
    class Categorical:
        def __init__(self, mapping):
            # check that the conversion is to valid colors
            assert(mcolors.is_color_like(color) for color in mapping.values())
            self._mapping = mapping

        def __call__(self, value):
            # convert value to a color
            return [mcolors.to_rgba(self._mapping[v]) for v in values]
    \end{minted}
    That we can test for action equivariance
    \begin{minted}{python}
    def test_nominal(values, encoder):
        m1 = list(zip(values, encoder(values)))
        random.shuffle(values)
        m2 = list(zip(values, encoder(values)))
        assert sorted(m1) == sorted(m2)
    \end{minted}
\end{frame}

\subsection{Artist}
\begin{frame}[fragile]{Artist}
    \begin{minted}{python}
        class ArtistClass(matplotlib.artist.Artist):
            def __init__(self, data, transforms, *args, **kwargs):
                # properties that are specific to the graphic
                self.data = data 
                self.transforms = transforms
                super().__init__(*args, **kwargs)

@@

\begin{frame}[fragile]{Artists: Scatter \& Line}
\begin{minted}{python}
class Point(mcollections.Collection):
    def assemble(self, x, y, s, facecolors='C0' ):
        # construct geometries of the circle glyphs in visual coordinates
        self._paths = [mpath.Path.circle(center=(xi,yi), radius=si) 
                    for (xi, yi, si) in zip(x, y, s)] 
        # set attributes of glyphs, these are vectorized 
        # circles and facecolors are lists of the same size
        self.set_facecolors(facecolors)

class Line(mcollections.LineCollection):
    def assemble(self, x, y, color='C0'):
        #assemble line marks as set of segments 
        segments = [np.vstack((vx, vy)).T for vx, vy in zip(x, y)]
        self.set_segments(segments)
        self.set_color(color)
\end{minted}
\end{frame}
\subsection{Data}
\begin{frame}[fragile]{Continuity}
\begin{minted}{python}
class VertexSimplex: #maybe change name to something else
    """Fiberbundle is consistent across all sections
    """
    FB = FiberBundle({'tables': ['vertex']},  
            {'v1': float, 'v2': str, 'v3': float})
    def __init__(self, sid = 45, size=1000, max_key=10**10):
        # create random list of keys
    def tau(self, k):
        # e1 is sampled from F1, e2 from F2, etc...
        return (k, (e1, e2, e3, e4))

class EdgeSimplex: 
    FB = FiberBundle({'tables': ['vertex','edge']},
                     {'x' : float, 'y':  float, 'color':mtypes.Color()}})
    def tau(self, k): #will fix location on page on revision
        x, y = self._xy(k, self.distances, 
                        self.angle_samples[k], self.angle_samples[k+1]) 
        color = self._color(k) 
        return (k, (x, y, color))
\end{minted}
\end{frame}

@@

@@

\begin{frame}{Glyph}
    The glyph is the graphic generated by $\vmark(\gbase_{\dbasepathpoint})$ where the path connected components $\dbasepath \subset \dbase$ are defined 
    \begin{equation}
    \dbasepath = \{\dbasepathpoint \in \dbase \text{ s. t. } \exists \gamma \text{ s.t. } \gamma(0)=\dbasepoint \text{ and }\gamma(1)=\dbasepathpoint\}
    \end{equation}
    such that the  path $\gamma$ from \dbasepoint\ to \dbasepathpoint\ is a continuous function from the interval [0,1] and $\gbase_{\dbasepathpoint}$ is the region
    \begin{equation}
        \begin{tikzcd}[ampersand replacement=\&]
            \gtotal \arrow[r, shift left] \& \gbase_\dbasepathpoint \arrow[rr, "\vindex(\gbasepoint)", shift left] \arrow[l, "\gsection(\gbase_\dbasepathpoint)"] \& \& \dbasepath_{\dbasepoint} \arrow[ll, "\vindex^{-1}(\dbasepath)"]
            \end{tikzcd}
        \label{eq:mark}
    \end{equation}
    such that the glyph is differentiable, in keeping with Ziemkiewicz and Kosara's description of a glyph\cite{ziemkiewiczEmbeddingInformationVisualization2009}.
\end{frame}
\begin{frame}[fragile]{Fiber Bundle}
    \begin{minted}{python}
    @dataclass
    class FiberBundle:
    """
    Attributes
    ----------
    K: {'tables': []}
    F: {variable name: type}
    """
        K: dict 
        F: dict
    \end{minted}
\end{frame}
